\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2018-11-14 draft.tex
polish

MapReduce is one of the most popular framework for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling polices and network latency, the same inputs may arrive at a reducer in different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the outputs of a reducer is independent of the order of its inputs. 
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier then the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

@@

On the other hand, in many cases, the non-commutative behavior of a reducer can be the source of some very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to a SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach working on real world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

@@

\textbf{The Prover component} reduces the commutativity problem to a SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to a SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative. 
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
We assume the reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutative of the produced reducer implies the commutative of the original one, but not vice verse.

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;3;4;5] and any other input with two consecutive values being swapped, e.g., [1;2;\textbf{4};\textbf{3};5].
The swap operation can be applied arbitrarily many times to produce all possible permutations of the input. It follows that that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;\textbf{3};\textbf{4};5], [1;\textbf{2};\textbf{4};3;5], [\textbf{1};\textbf{4};2;3;5], [4;1;2;3;5], etc. 
Recall that the first program transformation task stores all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

@@

\section{Optimizations}
\label{section:optimizations}
In this section we explain how data flow analysis is used in J-Recover to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

@@

\begin{figure}[hbt]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$s:= 0$; \;
			\Loop{}{
				$t:=\cur$;
				$s:= s + t$\;
			};
			$out(s)$\;\;
		\end{algorithm}
		\caption{A commutative reducer with an invalid Equation~(\ref{eq:commu}).}
		\label{fig:reducer_opt}
	\end{minipage}
	\begin{minipage}{0.55\textwidth}
		\LinesNumbered
		\centering
		\begin{minipage}{0.75\textwidth}
		\begin{algorithm}[H]
			$c_1;c_2;\ldots;c_i$\;
			\lIf{$g_1$}{$\ldots$}
			\lElseIf{ $g_2$}{$\ldots$}
			\lElseIf{ $g_3$}{$\ldots$}
			\lElseIf{ $g_m$}{$c_{i+1};\ldots ;c_n; out(v);\ldots$}
			\lElse{ $\ldots$}
		\end{algorithm}
	\end{minipage}
		\caption{$out(v)$ in nested branches of $s_3$.}
		\label{fig:nested_out}
	\end{minipage}
\end{figure}

@@

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember important variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$. 
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
\begin{itemize}
	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$. 
	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
\end{itemize}
After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of equation~(\ref{eq:commu2}) is critical to the success of our approach. In our evaluation, the ratio of examples that our approach can successfully analyze is significantly increased from $6.8\%$ to $97.5\%$. The details of the evaluation can be found in Section~\ref{section:exp}.

@@

The computation of $F$ in equation~(\ref{eq:commu2}) can be done in several different ways. For example, one can construct a formula in the style of bounded model checking. Here we apply \emph{symbolic execution} to generate $F$. Below we first review the standard symbolic execution algorithm and then explain how we use it to construct the formula $F$.

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer. 
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $(v:= e)$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $(v:= e)$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$. 
\end{enumerate}

@@

For the loop body $s_2$, symbolic execution is guaranteed to terminate because there is no commands in $s_2$ that may lead to infinite computations.
When the execution terminates, we collect all states with empty commands $\langle pc, val, \emptyset \rangle$ as a set $T$ and use them to create the function $F$ in Equation~(\ref{eq:commu2}). 
First, we create for each variable $x\in\Var$ a formula $f_x$ whose value is $\bigvee_{\langle pc, val, \emptyset \rangle \in T} (pc \wedge val(x))$.
The function $F(n,x_1,x_2,\ldots,x_k)$ simply returns the tuple $\{f_{x_1},f_{x_2},\ldots, f_{x_k}\}$.

@@

\begin{itemize}
	\item For the case that $\circ \in\{+,-\}$, we have $sv'_x(z) = sv_p(z)\circ  sv_q(z)$.
	\item For the case that $\circ \notin\{+,-\}$, we define $sv'_x(val(p) \circ val(q)) =1$ and $sv'_x(e) = 0$ for all $e\neq val(p) \circ val(q)$.
\end{itemize}
To make the definition complete for our simple reducer program commands, we further define (1) $sv_{\cur}(c_0) =1$ and $sv_{\cur}(e)=0$ for $e\neq \cur$, (2) $sv_{n}(1) =n$ and $sv_{n}(e)=0$ for $e\neq 1 \wedge n\in \Z$.
This data structure is efficient for strong symbolic values of reducer programs mainly for reason that the operations $+$ and $-$ are quite heavily used in such programs. 
Here we give a tiny example. Initially, $sv_x(x_0)=1$. After the commands $x:=x*x;x:=x+x$, we get $sv_x(x_0*x_0)=1$ and $sv_x(x_0*x_0)=2$, respectively.

\section{Evaluation}
\label{section:exp}
J-ReCoVer is implemented in Java and built on top of Soot~\cite{soot} 2.5.0 and Z3~\cite{z3} 4.7.1. We run J-Recover on a virtual machine running on a server with AMD Opteron 6376 CPU. 4GB of memory is allocated to this virtual machine. The operation system is Ubuntu 16.04.5 LTS.

In order to properly evaluate the performance of J-ReCoVer, we collect benchmark example from two different sources. 
First, we use the search engine \url{http://searchcode.com}~\cite{searchcode} to collect java programs containing the key strings ``public void reduce'' or ``protected void reduce''. Since there is an upper bound on number of results returned from the search engine, we add different search filters in order to get more data. We tried all $12$ combinations of six filters on code length $\{<50$, $ 50\sim 250$, $250\sim 450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$ 
and two filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total we get $11346$ Java programs. We exclude cases that are not Hadoop MapReduce reducer programs (those do not import the Hadoop library, do not extend or implement the reducer interface) and obtain $1273$ examples. We further remove those that cannot be compiled, those that are identical to each other, or those uses non-numerical data types (e.g, string) and obtained $118$ programs as the final benchmark.
The second set of examples are those have been discussed in verification literature~\cite{ChenHSW15,ChenSW16}. This set is relatively small, we have only $10$ examples. 
We also created a program that randomly generate reducer program in order to test the scalability of J-Recover.
All benchmark examples can be found on the web-site of J-Recover.

Our evaluation answers the following two questions:
\begin{itemize}
	\item Is J-Recover better than other competing tools?
	\item Are the two optimizations proposed in Section~\ref{section:optimizations} critical and effective?
\end{itemize}

\subsection{Comparison with competing tools}
As already pointed out in~\cite{ChenHSW15}, a reducer commutativity problem can be reduced to a reachability problem of a program. 
The examples in~\cite{ChenHSW15}, which includes \textsf{rangesum}\footnote{The sum of the second half of the input list.}, \textsf{avg}, \textsf{max}, \textsf{sep}\footnote{The number of occurrences of even numbers minus that of odd numbers.}, \textsf{sum}, are already included in SV-Comp~\cite{svcomp} (Software Verification Competition) benchmarks.
Although those examples are tiny and simple, none of the software verifiers can verify all of those examples according to the 2018 results. In contrast, J-Recover can verify all them very efficiently.

In Table~\ref{tab:svcomp}, the statistics of tools other than J-Recover are taken directly from the SV-Comp~\cite{svcomp} web-site. Although the machine we used are different, the results of J-Recover will not be much different if we run it on other machines. The other limitation is that in J-Recover, we analyze Java implementations and in SV-Comp they analyze the C implementations of the same functions. Since those functions to be analyzed are relatively simple, the syntactical difference between C and Java implementation is very small. The difference in the statistics mainly comes from the verification algorithm being used, not the programming language the reducer is written in.
\begin{table}[t]
\scalebox{0.95}{
	\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
		& \small CBMC\cite{cbmc2} 	& \small CPA-Seq\cite{cpachecker} & \small  ESBMC-kind\cite{esbmc} & \small  UAutomizer\cite{uautomizer} &\small  VeriAbs\cite{veriabs} &\small J-Recover\\
\hline
\hline
\small \textsf{rangesum}	& 0.96 	   & 30 		 & 2.2 				 &  10				 & 180       &9.9\\
\hline
\small \textsf{avg}			 &  TO      &  TO		  &  TO				 &  TO				 &  2          &10.9\\
\hline
\small \textsf{max}		    &  TO      &  TO		 &  TO				& TO				 &  7         &9.4\\
\hline
\small \textsf{sep}		     &  TO      &  TO		  &  TO				 & TO				 &  TO        &9.9\\
\hline
\small \textsf{sum}		    &  TO      &  TO		 &  TO				& TO				&  1.7       &9.9\\
\hline
	\end{tabular}
}
	\caption{SV-Comp 2018 results on reducer commutativity. For space reason, we only list results of competitive tool in this table, which includes top three tools in the overall ranking and the reachability safety category.}\vspace{-1cm}
	\label{tab:svcomp}
\end{table}

\subsection{The performance of J-Recover with and without optimizations}

For the optimization based on data flow analysis (Section~\ref{section:opt1}), we evaluation them based on real-world benchmark sets collected from open repositories (118 examples) and also literature (10 examples). The difference between the version with and without this optimization is very clear. The results are presented in table~\ref{tab:opt1}. The first column is the result without optimization, i.e., using Equation~(\ref{eq:commu}) and the second column is the result with optimization, i.e., using Equation~(\ref{eq:commu2}). The precision of the analysis is improved from $6.8\%$ to $97.5\%$ for examples from open repositories and
$10\%$ to $100\%$ for examples from literature. There are only three cases that J-Recover with data flow analysis cannot handle. The reason that J-Recover fails is that the three reducers use more complicated structure than what J-Recover currently supports; they use a branch statement before entering the loop, i.e., in the form $\ite{g}{(s_1;\rloop\{s_2\};s_3)}{(s'_1;\rloop\{s'_2\};s'_3)}$. In theory, such program can be handled by more sophisticated program transformation operations. For example, we can merge the two loops and push the outer branch condition inside the merged loop. 
However, this new program transformation task is not implemented in J-Recover yet.

>>>>

MapReduce is one of the most popular framework for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling polices and network latency, the same inputs may arrive at a reducer in different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the output of a reducer is independent of the order of its inputs. 
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier then the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

@@

On the other hand, in many cases, the non-commutative behavior of a reducer is the source of very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to an SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach working on real world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

@@

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative. 
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
We assume the reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutativity of the produced reducer implies the commutativity of the original one, but not vice verse.

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;\textbf{3};\textbf{4};5] and any other input with two consecutive values being swapped, e.g., [1;2;4;3;5].
The swap operation can be applied arbitrarily many times to produce all possible permutations of the input.
It follows that that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;3;4;5], [1;2;4;3;5], [1;4;2;3;5], [4;1;2;3;5], etc. 
Recall that the first program transformation task stores all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

@@

\section{Optimizations}
\label{section:optimizations}
In this section we explain how data flow analysis is used in J-ReCoVer to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

@@

\begin{figure}[hbt]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$s:= 0$; \;
			\Loop{}{
				$t:=\cur$;
				$s:= s + t$\;
			};
			$out(s)$\;\;\;
		\end{algorithm}
		\caption{A commutative reducer with an invalid Equation~(\ref{eq:commu}).}
		\label{fig:reducer_opt}
	\end{minipage}
	\ \ \ \ \ \ \ \ 
	\begin{minipage}{0.5\textwidth}
		\LinesNumbered
		\centering
		\begin{minipage}{0.75\textwidth}
		\begin{algorithm}[H]
			$c_1;c_2;\ldots;c_i$\;
			\lIf{$g_1$}{$\ldots$}
			\lElseIf{ $g_2$}{$\ldots$}
			\lElseIf{ $g_3$}{$\ldots$}
			\lElseIf{ $g_m$}{$c_{i+1};\ldots ;c_n; out(v);\ldots$}
			\lElse{ $\ldots$}
		\end{algorithm}
	\end{minipage}
		\caption{The $s_3$ command with an $out(v)$ command in nested branches.}
		\label{fig:nested_out}
	\end{minipage}
\end{figure}

@@

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember the variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$. 
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
\begin{itemize}
	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$. 
	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
\end{itemize}
After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of Equation~(\ref{eq:commu2}) is critical to the success of our approach. In our evaluation, the ratio of examples that our approach can successfully analyze is significantly increased from $6.8\%$ to $97.5\%$. The details of the evaluation can be found in Section~\ref{section:exp}.

@@

The computation of $F$ in Equation~(\ref{eq:commu2}) can be done in several different ways. For example, one can construct a formula in the style of bounded model checking. Here we apply \emph{symbolic execution} to generate $F$. Below we first review the standard symbolic execution algorithm and then explain how we use it to construct the formula $F$.

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer. 
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $v:= e$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $v:= e$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$. 
\end{enumerate}

@@

For the loop body $s_2$, symbolic execution is guaranteed to terminate because there is no commands in $s_2$ that may lead to infinite computations.
When the execution terminates, we collect all states with empty commands $\langle pc, val, \emptyset \rangle$ as a set $T$ and use them to create the function $F$ in Equation~(\ref{eq:commu2}). 
First, we create for each variable $x\in\Var$ a formula $f_x$ whose value is $\bigvee_{\langle pc, val, \emptyset \rangle \in T} (pc \wedge val(x))$.
The function $F(n,x_1,x_2,\ldots,x_k)$ simply returns the tuple $(f_{x_1},f_{x_2},\ldots, f_{x_k})$.

@@

\begin{itemize}
	\item For the case that $\circ \in\{+,-\}$, we have $sv'_x(z) = sv_p(z)\circ  sv_q(z)$.
	\item For the case that $\circ \notin\{+,-\}$, we define $sv'_x(val(p) \circ val(q)) =1$ and $sv'_x(e) = 0$ for all $e\neq val(p) \circ val(q)$.
\end{itemize}
To make the definition complete for our simple reducer program commands, we further define (I) $sv_{\cur}(c_0) =1$ and $sv_{\cur}(e)=0$ for $e\neq \cur$, (II) $sv_{n}(1) =n$ and $sv_{n}(e)=0$ for $e\neq 1 \wedge n\in \Z$.
This data structure is efficient for strong symbolic values of reducer programs mainly for reason that the operations $+$ and $-$ are quite heavily used in such programs. 
Here we give a tiny example. Initially, $sv_x(x_0)=1$. After the commands $x:=x*x;x:=x+x$, we get $sv_x(x_0*x_0)=1$ and $sv_x(x_0*x_0)=2$, respectively.

\section{Evaluation}
\label{section:exp}
J-ReCoVer is implemented in Java and built on top of Soot~\cite{soot} 2.5.0 and Z3~\cite{z3} 4.7.1. We run J-ReCoVer on a virtual machine running on a server with AMD Opteron 6376 CPU. 4GB of memory is allocated to this virtual machine. The operation system is Ubuntu 16.04.5 LTS.

In order to properly evaluate the performance of J-ReCoVer, we collect benchmark example from two different sources. 
First, we use the search engine \emph{searchcode.com}~\cite{searchcode} to collect java programs containing the key strings ``public void reduce'' or ``protected void reduce''. Since there is an upper bound on number of results returned from the search engine, we add different search filters in order to get more data. We tried all $12$ combinations of six filters on code length $\{<50$, $ 50\sim 250$, $250\sim 450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$ 
and two filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total we get $11346$ Java programs. We exclude cases that are not Hadoop MapReduce reducer programs (those do not import the Hadoop library, do not extend or implement the reducer interface) and obtain $1273$ examples. We further remove those that cannot be compiled, those that are identical to each other, or those with non-numerical data types (e.g, string) and obtained $118$ programs as the final benchmarks.
The second set of examples are those have been discussed in verification literature~\cite{ChenHSW15,ChenSW16}. This set is relatively small, we have only $10$ examples. 
We also created a program that randomly generate reducer program in order to test the scalability of J-ReCoVer.
All benchmark examples can be found on the web-site of J-ReCoVer.

Our evaluation answers the following two questions:
\begin{itemize}
	\item Is J-ReCoVer better than other competing tools?
	\item Are the two optimizations proposed in Section~\ref{section:optimizations} critical and effective?
\end{itemize}

\subsection{Comparison with competing tools}
As already pointed out in~\cite{ChenHSW15}, a reducer commutativity problem can be reduced to a reachability problem of a program. 
The examples in~\cite{ChenHSW15}, which includes \textsf{rangesum}\footnote{The sum of the second half of the input list.}, \textsf{avg}, \textsf{max}, \textsf{sep}\footnote{The number of occurrences of even numbers minus that of odd numbers.}, \textsf{sum}, are already included in SV-Comp~\cite{svcomp} (Software Verification Competition) benchmarks.
Although those examples are tiny and simple, none of the software verifiers can verify all of those examples according to the 2018 results. In contrast, J-ReCoVer can verify all of them very efficiently.

In Table~\ref{tab:svcomp}, the statistics of tools other than J-ReCoVer are taken directly from the SV-Comp~\cite{svcomp} web-site. Although the machine we used are different, the results of J-ReCoVer will not be much different if we run it on other machines. The other issue of this experiment is that in J-ReCoVer, we analyze the Java implementations and in SV-Comp they analyze the C implementations of the same functions. Since those functions to be analyzed are relatively simple, the syntactical difference between C and Java implementation is very small. The difference in the statistics mainly comes from the verification algorithm being used, not the programming language the reducer is written in.
\begin{table}[t]
\scalebox{0.95}{
	\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
		& \small CBMC\cite{cbmc2} 	& \small CPA-Seq\cite{cpachecker} & \small  ESBMC-kind\cite{esbmc} & \small  UAutomizer\cite{uautomizer} &\small  VeriAbs\cite{veriabs} &\small J-ReCoVer\\
\hline
\hline
\small \textsf{rangesum}	& 0.96 	   & 30 		 & 2.2 				 &  10				 & 180       &9.9\\
\hline
\small \textsf{avg}			 &  TO      &  TO		  &  TO				 &  TO				 &  2          &10.9\\
\hline
\small \textsf{max}		    &  TO      &  TO		 &  TO				& TO				 &  7         &9.4\\
\hline
\small \textsf{sep}		     &  TO      &  TO		  &  TO				 & TO				 &  TO        &9.9\\
\hline
\small \textsf{sum}		    &  TO      &  TO		 &  TO				& TO				&  1.7       &9.9\\
\hline
	\end{tabular}
}
	\caption{SV-Comp 2018 results on reducer commutativity. For space reason, we only list results of competitive tool in this table, which includes top three tools in the overall ranking and the reachability safety category.}\vspace{-1cm}
	\label{tab:svcomp}
\end{table}

\subsection{The performance of J-ReCoVer with and without optimizations}

For the optimization based on data flow analysis (Section~\ref{section:opt1}), we evaluate them based on real-world benchmark sets collected from open repositories (118 examples) and also literature (10 examples). The difference between the version with and without this optimization is very clear. The results are presented in table~\ref{tab:opt1}. The first column is the result without optimization, i.e., using Equation~(\ref{eq:commu}) and the second column is the result with optimization, i.e., using Equation~(\ref{eq:commu2}). The precision of the analysis is improved from $6.8\%$ to $97.5\%$ for examples from open repositories and
$10\%$ to $100\%$ for examples from literature. There are only three cases that J-ReCoVer with data flow analysis cannot handle. The reason that J-ReCoVer fails is that the three reducers use more complicated structure than what J-ReCoVer currently supports; they use a branch statement before entering the loop, i.e., in the form $\ite{g}{(s_1;\rloop\{s_2\};s_3)}{(s'_1;\rloop\{s'_2\};s'_3)}$. In theory, such program can be handled by more sophisticated program transformation tasks. For example, we can merge the two loops and push the outer branch condition into the merged loop. 
However, this new program transformation task is not implemented in J-ReCoVer yet.

If we look into scrutiny, we failed to verify any example using Equation~(\ref{eq:commu}). The reason is that when Soot convert a Java program to the Jimple format, it always uses a temporary variable to store the value of $\cur$ in the loop body (similar to the example in Figure~\ref{fig:reducer_opt}). We tried a simpler heuristics approach that only ignores the temporary variables and the result is not as good as using data flow analysis, because the values of the temporary variables often also propagate to other variables. 
\begin{table}[htb]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		& &W/O optimization	& With optimization\\
		\hline
		\hline
		\multirow{4}{*}{Open repositories}&Commutative& 0&106\\
		\cline{2-4}
		&Non-commutative&8&9\\
		\cline{2-4}
		&Unknown&110&3\\
		\cline{2-4}
		&Precision& 6.8\% & 97.5\%\\
		\hline
		\hline
		\multirow{4}{*}{Literatures}&Commutative& 0&9\\
		\cline{2-4}
		&Non-commutative&1&1\\
		\cline{2-4}
		&Unknown&9&0\\
		\cline{2-4}
		&Precision& 10\% & 100\%\\
		\hline
	\end{tabular}
	\caption{The improvement of precision using data flow analysis. The row ``Precision'' is calculated as the number of examples that J-ReCoVer returns commutative or non-commutative divided by the total number of example.}
	\label{tab:opt1}
\end{table}
